---
title: "Using createGRanges with karyoploteR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using createGRanges with karyoploteR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pidProjCode)
library(karyoploteR)
library(data.table)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
```

As a sample data set we use a subset of the data from a GWAS of selective IgA deficiency by Bronson et al. (2016). This contains only the SNPs on chromosome 6 from 25Mb to 36Mb. 

```{r loadExampleData}
#dat<-fread(file=system.file('inst/extdata', 'karyoploterExampleData.tsv.gz', package='pidProjCode', mustWork=T), sep='\t', header=T, select=c('SNPID','CHR38','BP38','P'))

dat<-fread(file='karyoploterExampleData.tsv.gz', sep='\t', header=T, select=c('SNPID','CHR38','BP38','P'))
```

The first obstacle to getting `karyoploteR` working was getting GWAS data into a `GenomicRanges` object. I found the function `cpvSnp::createArrayData` which achieved something like this and wrote `createGRanges` as a stripped-down version of it. 

To create a `GenomicRanges` object, we need to specify the data and the names of the columns giving the chromosome and basepair coordinates. Any other columns will be kicked into the metadata. As written, `createGRanges` cannot take a `data.table` object so we need to convert `dat` to a `data.frame` first.

A useful feature of `karyoploteR` is the highlighting of specified SNPs. These need to be given in a `GenomicRanges` object, too, so we create `topSnpGRanges` to hold a single SNP. For a Manhattan plot we want a y axis on the negative log10 scale, but `karyoploteR` doesn't seem to heed this for highlighted SNPs, so we specify the y coordinate as an extra metadata column in the `GenomicRanges` object

```{r creatingGRangesObjects}
gRanges<-createGRanges(data.frame(dat), bpCol='BP38', chrCol='CHR38')

topSnps<-c('rs116540075')

topSnpGRanges<-createGRanges(data.frame(dat[SNPID %in% topSnps][, y:= -log10(P)]), bpCol='BP38', chrCol='CHR38')
```

`createGRanges` is the only function from this package (`pidProjCode`) called in this vignette. The rest come from `karyoploteR` or its imports.

```{r manhattan}
ymax<-137
r0<-0

kp<-plotKaryotype(plot.type=4, labels.plotter=NULL, zoom="chr6:25e6-36e6")
title(main='IgAD GWAS, chr6:25M-36M, MHC boxed in blue, top extra-MHC SNP rs116540075 circled in red', cex.main=2.7)
kpAddBaseNumbers(kp, add.units=T, cex=1, tick.dist=5e5)
kpAxis(kp, ymin=0,ymax=ymax, r0=r0)
kp<-kpPlotManhattan(kp, data=gRanges, points.col='brewer.set3', ymax=ymax, r0=r0)
kpRect(kp, chr='chr6', x0=28510120, x1=33480577,r0=r0, y0=0,y1=1, col=NA, border='blue', lwd=3)
kp<-kpPoints(kp, data=topSnpGRanges, pch=1, cex=1.6, col="red", lwd=2, ymax=ymax, r0=r0)
```

```{r zoomedManhattan}
ymax<-38
r0<-0.2

kp<-plotKaryotype(plot.type=4, labels.plotter=NULL, zoom="chr6:33e6-34e6")
title(main='IgAD GWAS, chr6:33M-34M, neighbourhood of top extra-MHC SNP rs116540075 (circled in red)', cex.main=2.7)
kpAddBaseNumbers(kp, add.units=T, cex=1, tick.dist=1e5)
kpAxis(kp, ymin=0,ymax=ymax, r0=r0)
kp<-kpPlotManhattan(kp, data=gRanges, points.col='brewer.set3', ymax=ymax, r0=r0)
kpRect(kp, chr='chr6', x0=33e6, x1=33480577,r0=r0, y0=0,y1=1, col=NA, border='blue', lwd=3)
kp<-kpPoints(kp, data=topSnpGRanges, pch=1, cex=1.6, col="red", lwd=2, ymax=ymax, r0=r0)

genes.data<-makeGenesDataFromTxDb(karyoplot=kp, txdb=TxDb.Hsapiens.UCSC.hg38.knownGene)

genes.data<-addGeneNames(genes.data)
genes.data.merged<-mergeTranscripts(genes.data)
kp<-kpPlotGenes(kp, data=genes.data.merged, r0=0, r1=0.2, cex=1.0, gene.name.position='left')
```